The Semantics and Implementation of \CEU: a Synchronous Reactive Language based on Esterel

### 2015-09-24 paper.tex
Cenarios onde a restricao de "delayed await" atrapalha.

Another distinction regarding event handling is that Esterel supports 
same-cycle bi-directional communication~\cite{esterel.compiling}, i.e., two 
threads can react to one another during the same cycle due to mutual signal 
dependency.
%
\CEU imposes a restriction that an \code{await} is only valid for the next 
reaction, i.e., if an \code{await} and \code{emit} for the same event occur 
simultaneously in parallel trails, the \code{await} does not awake.
%
These \emph{delayed awaits} avoid corner cases of instantaneous termination and 
re-execution of statements in the same reaction inside loops (known as 
\emph{schizophrenic statements})~\cite{esterel.schizo}.
%
The example of Figure~\ref{lst.delay} relies on this restriction to avoid 
infinite execution:
Although both sides of the \code{par/or} have an \code{await} statement, the 
\code{loop} would restart instantaneously if the \code{await} could awake in 
the same reaction that reaches it.
%
In an atypical scenario requiring immediate awake, delayed awaits can be 
circumvented by judiciously copying the code to execute on awake.

\begin{figure}
\begin{lstlisting}
event void e,f;
loop do
    par/or do
        await e;
    with
        emit e;
        await f;
    end
end
\end{lstlisting}
\caption{ The \code{await} occurs simultaneously with the \code{emit} and does 
not awake, preventing the \code{par/or} to terminate and execute the 
\code{loop} indefinitely.
\label{lst.delay}
}
\end{figure}

>>>>

Another distinction regarding event handling is that Esterel supports 
same-cycle bi-directional communication~\cite{esterel.compiling}, i.e., two 
threads can react to one another during the same cycle due to mutual signal 
dependency.
%
\CEU imposes a restriction that an \code{await} is only valid for the next 
reaction, i.e., if an \code{await} and \code{emit} for the same event occur 
simultaneously in parallel trails, the \code{await} does not awake.
%
These \emph{delayed awaits} avoid corner cases of instantaneous termination and 
re-execution of statements in the same reaction inside loops (known as 
\emph{schizophrenic statements})~\cite{esterel.schizo}.
%
The example in the left of Figure~\ref{lst.delay} (\code{CODE-1}) relies on 
this restriction to avoid infinite execution:
Although both sides of the \code{par/or} have an \code{await} statement, the 
\code{loop} would restart instantaneously if the \code{await} could awake in 
the same reaction that reaches it.
%
In atypical scenarios requiring immediate awake, delayed awaits can be 
circumvented by manually copying or making some transformation in the code to 
execute on awake.
%
From our experience, in some cases we need to execute a block of code 
periodically, \emph{including the current reaction}, from requests of an 
internal event, as illustrated in the middle of Figure~\ref{lst.delay} 
(\code{CODE-2}).
In this case, moving the \code{await} to the end of the loop suffices to 
execute the periodic code immediately (and then in reaction to each \code{emit} 
request).
If the periodic \code{emit} depends on a condition, the code becomes more 
intricate, requiring to copy the condition test to the periodic code, as 
illustrated in the right of Figure~\ref{lst.delay} (\code{CODE-3}).

\begin{figure}
\begin{minipage}[t]{0.24\linewidth}
\begin{lstlisting}
event void e,f;
loop do
  par/or do
    await e;
  with
    emit e;
    await f;
  end
end

/* CODE-1 */
\end{lstlisting}
\end{minipage}
%
\begin{minipage}[t]{0.36\linewidth}
\begin{lstlisting}
event void e;
par do
  loop do
    <...>
    emit e;  // periodic request
  end
with
  loop do
    <...>    // code to execute
    await e; // await after
  end
end

/* CODE-2 */
\end{lstlisting}
\end{minipage}
%
\begin{minipage}[t]{0.36\linewidth}
\begin{lstlisting}
event void e;
var bool execute_now = false;
par do
  loop do
    <...>
    if <...> then
        execute_now = true;
        emit e;
    end
  end
with
  loop do
    if execute_now then
      <...> // code to execute
    end
    await e;
  end
end

/* CODE-3 */
\end{lstlisting}
\end{minipage}
%
\caption{ In \code{CODE-1}, the \code{await} occurs simultaneously with the 
\code{emit} and does not awake, preventing the \code{par/or} to terminate and 
execute the \code{loop} indefinitely.
\code{CODE-2} and \code{CODE-3} illustrate how to circumvent the \emph{delayed 
await} restriction.
\label{lst.delay}
}
\end{figure}
